---
title: ""
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: hide
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Project 1 Brief

In this project, we have a text file with chess tournament results where the information has some structure. The purpose of the project is to create a structured .csv file with below variables as columns. 

- Player's Name
- Player's State
- Total Number of Points
- Player's Pre Rating
- Average Pre Chess Rating of Opponents

The first information would be : Gary Hua, ON, 6.0, 1794, 1605

## Data Collection in txt format

```{r}
install.packages('stringr', repos = "http://cran.us.r-project.org")
library('stringr')
```



```{r}
# created a file variable and assign the txt url 
# header is false
#skipped first 4 rows as it is dont need them.

file <- 'https://raw.githubusercontent.com/anilak1978/chess-tournament/master/tournamentinfo.txt'
raw_data <- read.table(file, header = FALSE, sep = ',', skip = 4)
head(raw_data)

```

## Find the Player ID



```{r}
#1 or more digit numbers, look for the ones that ends with white space and |

player_id <- unlist(str_extract_all(unlist(raw_data), "\\d{1,}(?=\\s\\|)"))
player_id

```

## Find the Player Name


```{r}
# All letters, any word character and a space after that and match it two or more times. 

player_name <- unlist(str_extract_all(unlist(raw_data), "([[:alpha:]]+\\s){2,}"))
player_name

```

## Find the Player State


```{r}
# all letters two of them together that ends with white space and |, ends with space and look for 

player_state <- unlist(str_extract_all(unlist(raw_data), "([[:alpha:]]){2}\\s(?=\\|)"))
player_state


```

## Find Total Points


```{r}
# digit , period followed by digit. 

total_points <- unlist(str_extract_all(unlist(raw_data), "\\d\\.\\d"))
total_points

```

## Find Player's Pre Rating


```{r}
#3 or 4 digit number that starts and ends with sspace

players_pre_rating <- unlist(str_extract_all(unlist(raw_data), "[^->] \\d{3,4}(?=\\s)|[^->] \\d{3,4}[[:alpha:]][[:digit:]]|  \\d{3,4}[[:alpha:]][[:digit:]]"))
players_pre_rating <- str_replace_all(players_pre_rating, pattern = ":", replacement = "")
players_pre_rating <- str_replace_all(players_pre_rating, pattern = " ", replacement = "")
players_pre_rating <- str_replace_all(players_pre_rating, pattern = "P[[:digit:]]", replacement = "")
players_pre_rating

```

# Drop spaces in findings



```{r}
# drop white spaces in all findings

player_name <- str_trim(player_name, "both")
player_state <- str_trim(player_state, "both")
total_points <- str_trim(total_points, "both")
players_pre_rating <- str_trim(players_pre_rating, "both")

```

# Create a dataframe

```{r}

df <- data.frame(player_id, player_name, player_state, total_points, players_pre_rating)
head(df)


```

# Look at structure of the dataframe

```{r}

str(df)

```

We need to convert to numerical variables in order to find the Average Pre Chess Rating of Opponents.

## Convert to numerical variables

```{r}
# convert o numerical variables
df$player_id <- as.numeric(df$player_id)
df$total_points <- as.numeric(df$total_points)
df$players_pre_rating <- as.numeric(df$players_pre_rating)
str(df)

```

# Find Opponent of Each Player


```{r}
# one or more digit number that ends with |

options(max.print = 2000)
opponent_player <- unlist(str_extract_all(unlist(raw_data), "(\\d{1,}|[[:blank:]]{1})(?=\\|)"))
opponent_player[opponent_player==" "] <- NA
opponent_player <- as.numeric(opponent_player)
opponent_player

```

# Map each player's opponent to Pre Rating

```{r}

# total 7 rounds of games played

round_1 <- opponent_player[seq(4, length(opponent_player),10)]
round_1 <- round_1[seq(1, length(round_1), 2)]
round_2 <- opponent_player[seq(5, length(opponent_player),10)]
round_2 <- round_2[seq(1, length(round_2), 2)]
round_3 <- opponent_player[seq(6, length(opponent_player),10)]
round_3 <- round_3[seq(1, length(round_3), 2)]
round_4 <- opponent_player[seq(7, length(opponent_player),10)]
round_4 <- round_4[seq(1, length(round_4), 2)]
round_5 <- opponent_player[seq(8, length(opponent_player),10)]
round_5 <- round_5[seq(1, length(round_5), 2)]
round_6 <- opponent_player[seq(9, length(opponent_player),10)]
round_6 <- round_6[seq(1, length(round_6), 2)]
round_7 <- opponent_player[seq(10, length(opponent_player),10)]
round_7 <- round_7[seq(1, length(round_7), 2)]

rounds <- matrix(c(round_1, round_2, round_3, round_4, round_5, round_6, round_7), ncol = 7)
rounds

```

# Calculate Average Pre Chess Rating of Opponent

```{r}

avg_opponent_player <- 0

for (i in 1:(length(player_id))){
  avg_opponent_player[i] <- mean(as.numeric(players_pre_rating[rounds[i,]]), na.rm=T)
}


```

# Add the Average Pre Chess Rating of Opponent to the dataframe

```{r}

df$avg_opponent_player <- avg_opponent_player
head(df)


```


## Export the dataframe to .csv file


```{r}

write.csv(df, file = "chess_tournament.csv")

```



